---
title: "ST558 Homework 2"
subtitle: "Lee Bennett"
format: html
editor: visual
---

## Task 1: Vectors

First, we'll create vectors to contain the blood pressure data from before treatment and after treatment, labeling each measurement by subject number:

```{r}
pre_bp <- c( 120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)

patients<-paste("Subject", 1:20, sep = "_")

names(pre_bp)<-patients
names(post_bp)<-patients
```

Next, let's calculate the change in blood pressure after treatment and compute the mean change across all subjects (pre-treatment minus post-treatment):

```{r}
diff_bp <- pre_bp-post_bp
diff_bp

mean(diff_bp)
```

We see that the average change in systolic blood pressure is a decrease of 5.65 mm Hg.

Next, we'll create a subset of the subjects who experienced a decrease in blood pressure and compute the average decrease in this subset:

```{r}
bp_decrease<-which(diff_bp>0)
bp_decrease_vals<-diff_bp[bp_decrease]
bp_decrease_vals
mean(bp_decrease_vals)
```

We can see that if we restrict to only those subjects whose blood pressure improved after treatment, the average decrease is 10.6 mm Hg.

## Task 2: Data Frames

For this task, we'll create a data frame containing the blood pressure data for each patient:

```{r}
bp_data <- data.frame(patients,pre_bp,post_bp,diff_bp)
head(bp_data)
```

Let's select only the observations where the blood pressure change is negative:

```{r}
`[`(bp_data,bp_data$diff_bp<0,1:4,drop=FALSE)
```

Finally, we create a logical variable called `bp_normal` to flag those observations where the post-treatment blood pressure is less than 120 mm Hg. The final data frame appears below.

```{r}
bp_data$bp_normal <- bp_data$post_bp<120
knitr::kable(bp_data)
```

## Task 3: Lists

Next, we'll create a data frame similar to the one above containing data from a similar experiment where the intervention was placebo instead of an active treatment:

```{r}
patients_p <- paste("Subject", 1:10, sep = "_")
pre_bp_p <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_bp_p <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
diff_bp_p <- pre_bp_p-post_bp_p
bp_normal_p <- post_bp_p<120

bp_data_p <- data.frame(patients_p,pre_bp_p,post_bp_p,diff_bp_p,bp_normal_p)
bp_data_p
```

We can now create a list that contains the original data frame with active treatment and the new data frame from placebo intervention as the two elements, named `treatment` and `placebo`, respectively:

```{r}
study_list <- list(treatment=bp_data, placebo=bp_data_p)
study_list
```

If we would like to access the first element in this list, we can accomplish that in a few different ways:

**1.  Using an explicit subscript to return the list element:**

```{r}
study_list[1]
```

**2.  Using double brackets to return a data frame:**

```{r}
study_list[[1]]
```

**3.  Using the name of the list element:**

```{r}
study_list$treatment
```

Finally, we can access the `pre_bp` column of the second data frame using one line of code to return a vector:

```{r}
study_list$placebo$pre_bp
```

